HMM Model

### 2020-10-31 manuscript/hmm.Rmd
Somewhat working new hypothesis code

do_pp_checks <- FALSE
hmm_hypothesis_res_list <- list()

@@

serie_data <- wide %>%
  inner_join(data$patient_data, by = c("hospital_id", "patient_id")) %>% 
  mutate(.time = day, .serie = patient_id, .observed = breathing_s) %>%
  transform_serie_data()

serie_data_28 <- wide %>%
  right_join(crossing(data$patient_data, day = 0:28), by = c("hospital_id", "patient_id", "day")) %>%
  mutate(.time = day, .serie = patient_id, .observed = breathing_s) %>%
  transform_serie_data()

# Extend terminal states and took_XX data for serie_data_28
serie_data_28 <- serie_data_28 %>%
  group_by(.serie) %>%
  mutate(last_day = max(day[!is.na(.observed)]),
         last_state = .observed[day == last_day],
         .observed = if_else(last_state %in% c("Discharged", "Death") & is.na(.observed) & day > last_day,
                             last_state,
                             .observed)) %>%
  mutate(across(tidyselect::starts_with("took_"), ~ if_else(is.na(.x), as.integer(any(.x != 0, na.rm = TRUE)), .x))) %>%
  ungroup()

@@

```{r}
# counterfactual_took_hcq <- serie_data_28 %>% mutate(took_hcq = TRUE)
# counterfactual_no_hcq <- counterfactual_base %>% mutate(took_hcq = FALSE)
# 
# 
# 
# trt_group_m1_counterfactual_took_hcq <- trt_group_m1
# trt_group_m1_counterfactual_took_hcq$serie_data = counterfactual_took_hcq
# 
# trt_group_m1_counterfactual_took_hcq_epred_rect <- posterior_epred_rect(trt_group_m1_fit, newdata = trt_group_m1_counterfactual_took_hcq)
# 
# trt_group_m1_counterfactual_no_hcq <- trt_group_m1
# trt_group_m1_counterfactual_no_hcq$serie_data = counterfactual_no_hcq
# 
# trt_group_m1_counterfactual_no_hcq_epred_rect <- posterior_epred_rect(trt_group_m1_fit, newdata = trt_group_m1_counterfactual_no_hcq)

@@

```{r}
# draws <- tidybayes::tidy_draws(trt_group_m1_fit$brmsfit)
# hmm_hypothesis_res_list[["trt_group_m1_death"]] <- 
#   bayesian_hypothesis_res_from_draws(
#     draws = draws$b_took_hcqTRUE,
#     model = "HMM",
#     estimand = "log(HR)",
#     hypothesis = hypotheses$hcq_reduces_death,
#     adjusted = "supportive"
#   )
# 
# hmm_hypothesis_res_list[["trt_group_m1_discharged"]] <- 
#   bayesian_hypothesis_res_from_draws(
#     draws = draws$b_took_hcqTRUE + draws$`b_rate_groupimprove_one:took_hcqTRUE`,
#     model = "HMM",
#     estimand = "log(HR)",
#     hypothesis = hypotheses$hcq_increases_discharged,
#     adjusted = "supportive"
#   )
```

```{r}
# fit <- trt_group_m1_fit
# epred_rect <- posterior_epred_rect(fit)
# predicted_rect <- posterior_epred_to_predicted(fit, epred_rect)
# predicted <- posterior_rect_to_long(fit, predicted_rect)

if(do_pp_checks) {
  do_common_pp_checks(trt_group_m1_fit)
}
```

@@

```{r}
# draws <- tidybayes::tidy_draws(trt_group_m2_fit$brmsfit)
# hmm_hypothesis_res_list[["trt_group_m2_death"]] <- 
#   bayesian_hypothesis_res_from_draws(
#     draws = draws$b_took_hcqTRUE,
#     model = "HMM",
#     estimand = "log(HR)",
#     hypothesis = hypotheses$hcq_reduces_death,
#     adjusted = "age, sex, supportive"
#   )
# 
# hmm_hypothesis_res_list[["trt_group_m2_discharged"]] <- 
#   bayesian_hypothesis_res_from_draws(
#     draws = draws$b_took_hcqTRUE + draws$`b_rate_groupimprove_one:took_hcqTRUE`,
#     model = "HMM",
#     estimand = "log(HR)",
#     hypothesis = hypotheses$hcq_increases_discharged,
#     adjusted = "age, sex, supportive"
#   )
```

@@

```{r}
# draws <- tidybayes::tidy_draws(trt_group_m3_fit$brmsfit)
# hmm_hypothesis_res_list[["trt_group_m3_death"]] <- 
#   bayesian_hypothesis_res_from_draws(
#     draws = draws$b_took_hcqTRUE,
#     model = "HMM",
#     estimand = "log(HR)",
#     hypothesis = hypotheses$hcq_reduces_death,
#     adjusted = "age, sex, supportive, hospital"
#   )
# 
# hmm_hypothesis_res_list[["trt_group_m3_discharged"]] <- 
#   bayesian_hypothesis_res_from_draws(
#     draws = draws$b_took_hcqTRUE + draws$`b_rate_groupimprove_one:took_hcqTRUE`,
#     model = "HMM",
#     estimand = "log(HR)",
#     hypothesis = hypotheses$hcq_increases_discharged,
#     adjusted = "age, sex, supportive, hospital"
#   )
```

@@

```{r}
# draws <- tidybayes::tidy_draws(trt_group_m4_fit$brmsfit)
# hmm_hypothesis_res_list[["trt_group_m4_death"]] <- 
#   bayesian_hypothesis_res_from_draws(
#     draws = draws$b_took_hcqTRUE,
#     model = "HMM",
#     estimand = "log(HR)",
#     hypothesis = hypotheses$hcq_reduces_death,
#     adjusted = "age, sex, supportive, hospital, comorbidities_sum"
#   )
# 
# hmm_hypothesis_res_list[["trt_group_m4_discharged"]] <- 
#   bayesian_hypothesis_res_from_draws(
#     draws = draws$b_took_hcqTRUE + draws$`b_rate_groupimprove_one:took_hcqTRUE`,
#     model = "HMM",
#     estimand = "log(HR)",
#     hypothesis = hypotheses$hcq_increases_discharged,
#     adjusted = "age, sex, supportive, hospital, comorbidities_sum"
#   )
```

@@

```{r}

# counterfactual_max_days <- 90
# counterfactual_base <-  crossing(day = 0:counterfactual_max_days, patient_id = unique(serie_data$patient_id)) %>%
#   inner_join(data$patient_data, by = c("patient_id")) %>% 
#   left_join(wide, by = c("hospital_id", "patient_id", "day")) %>%
#   mutate(.time = day, .serie = patient_id, .observed = breathing) %>%
#   mutate(.time = .time +1) %>% 
#   filter(.time >= 1) %>%
#   group_by(.serie) %>%
#   mutate(best_supportive_care = !is.na(best_supportive_care_from) & day >= best_supportive_care_from) %>%
#   ungroup()
# 
# 
# counterfactual_took_hcq <- counterfactual_base %>% mutate(took_hcq = TRUE)
# counterfactual_no_hcq <- counterfactual_base %>% mutate(took_hcq = FALSE)
# 
# 
# 
# trt_rate_m1_counterfactual_took_hcq <- trt_rate_m1
# trt_rate_m1_counterfactual_took_hcq$serie_data = counterfactual_took_hcq
# 
# trt_rate_m1_counterfactual_took_hcq_epred_rect <- posterior_epred_rect(trt_rate_m1_fit, newdata = trt_rate_m1_counterfactual_took_hcq)
# 
# trt_rate_m1_counterfactual_no_hcq <- trt_rate_m1
# trt_rate_m1_counterfactual_no_hcq$serie_data = counterfactual_no_hcq
# 
# trt_rate_m1_counterfactual_no_hcq_epred_rect <- posterior_epred_rect(trt_rate_m1_fit, newdata = trt_rate_m1_counterfactual_no_hcq)

```{r}
# draws <- tidybayes::tidy_draws(trt_rate_m1_fit$brmsfit)
# hmm_hypothesis_res_list[["trt_rate_m1_death"]] <- 
#   bayesian_hypothesis_res_from_draws(
#     draws = draws$b_took_hcqTRUE,
#     model = "HMM",
#     estimand = "log(HR)",
#     hypothesis = hypotheses$hcq_reduces_death,
#     adjusted = "supportive"
#   )
# 
# hmm_hypothesis_res_list[["trt_rate_m1_discharged"]] <- 
#   bayesian_hypothesis_res_from_draws(
#     draws = draws$b_took_hcqTRUE + draws$`b_rate_groupimprove_one:took_hcqTRUE`,
#     model = "HMM",
#     estimand = "log(HR)",
#     hypothesis = hypotheses$hcq_increases_discharged,
#     adjusted = "supportive"
#   )
```

@@

```{r}
# draws <- tidybayes::tidy_draws(trt_rate_m2_fit$brmsfit)
# hmm_hypothesis_res_list[["trt_rate_m2_death"]] <- 
#   bayesian_hypothesis_res_from_draws(
#     draws = draws$b_took_hcqTRUE,
#     model = "HMM",
#     estimand = "log(HR)",
#     hypothesis = hypotheses$hcq_reduces_death,
#     adjusted = "age, sex, supportive"
#   )
# 
# hmm_hypothesis_res_list[["trt_rate_m2_discharged"]] <- 
#   bayesian_hypothesis_res_from_draws(
#     draws = draws$b_took_hcqTRUE + draws$`b_rate_groupimprove_one:took_hcqTRUE`,
#     model = "HMM",
#     estimand = "log(HR)",
#     hypothesis = hypotheses$hcq_increases_discharged,
#     adjusted = "age, sex, supportive"
#   )
```

@@

```{r}
# draws <- tidybayes::tidy_draws(trt_rate_m3_fit$brmsfit)
# hmm_hypothesis_res_list[["trt_rate_m3_death"]] <- 
#   bayesian_hypothesis_res_from_draws(
#     draws = draws$b_took_hcqTRUE,
#     model = "HMM",
#     estimand = "log(HR)",
#     hypothesis = hypotheses$hcq_reduces_death,
#     adjusted = "age, sex, supportive, hospital"
#   )
# 
# hmm_hypothesis_res_list[["trt_rate_m3_discharged"]] <- 
#   bayesian_hypothesis_res_from_draws(
#     draws = draws$b_took_hcqTRUE + draws$`b_rate_groupimprove_one:took_hcqTRUE`,
#     model = "HMM",
#     estimand = "log(HR)",
#     hypothesis = hypotheses$hcq_increases_discharged,
#     adjusted = "age, sex, supportive, hospital"
#   )
```

@@

```{r}
# draws <- tidybayes::tidy_draws(trt_rate_m4_fit$brmsfit)
# hmm_hypothesis_res_list[["trt_rate_m4_death"]] <- 
#   bayesian_hypothesis_res_from_draws(
#     draws = draws$b_took_hcqTRUE,
#     model = "HMM",
#     estimand = "log(HR)",
#     hypothesis = hypotheses$hcq_reduces_death,
#     adjusted = "age, sex, supportive, hospital, comorbidities_sum"
#   )
# 
# hmm_hypothesis_res_list[["trt_rate_m4_discharged"]] <- 
#   bayesian_hypothesis_res_from_draws(
#     draws = draws$b_took_hcqTRUE + draws$`b_rate_groupimprove_one:took_hcqTRUE`,
#     model = "HMM",
#     estimand = "log(HR)",
#     hypothesis = hypotheses$hcq_increases_discharged,
#     adjusted = "age, sex, supportive, hospital, comorbidities_sum"
#   )
```

@@

```{r}
# draws <- tidybayes::tidy_draws(trt_m6_fit$brmsfit)
# hmm_hypothesis_res_list[["trt_m6_d_dimer_death"]] <- 
#   bayesian_hypothesis_res_from_draws(
#     draws = draws$b_log_peak_d_dimer,
#     model = "HMM",
#     estimand = "log(HR)",
#     hypothesis = hypotheses$d_dimer_increases_death,
#     adjusted = "age, sex, hospital"
#   )
#

@@

@@

```{r}
# draws <- tidybayes::tidy_draws(trt_m7_fit$brmsfit)
# hmm_hypothesis_res_list[["trt_m7_IL_6_death"]] <- 
#   bayesian_hypothesis_res_from_draws(
#     draws = draws$b_log_peak_IL_6,
#     model = "HMM",
#     estimand = "log(HR)",
#     hypothesis = hypotheses$IL_6_increases_death,
#     adjusted = "age, sex, hospital"
#   )

@@

@@

```
```{r}
if(do_pp_checks) {
  do_common_pp_checks(c_trt_m1_fit)
}
```

@@

@@

@@

```{r}
# draws <- tidybayes::tidy_draws(c_trt_m4_fit$brmsfit)
# hmm_hypothesis_res_list[["c_trt_m4_death"]] <- 
#   bayesian_hypothesis_res_from_draws(
#     draws = draws$b_took_hcqTRUE,
#     model = "HMM",
#     estimand = "log(HR)",
#     hypothesis = hypotheses$hcq_reduces_death,
#     adjusted = "age, sex, supportive, hospital, comorbidities_sum"
#   )
# 
# hmm_hypothesis_res_list[["c_trt_m4_discharged"]] <- 
#   bayesian_hypothesis_res_from_draws(
#     draws = draws$b_took_hcqTRUE + draws$`b_rate_groupimprove_one:took_hcqTRUE`,
#     model = "HMM",
#     estimand = "log(HR)",
#     hypothesis = hypotheses$hcq_increases_discharged,
#     adjusted = "age, sex, supportive, hospital, comorbidities_sum"
#   )
```

@@

```{r}
hmm_hypothesis_res_all <- do.call(rbind, hmm_hypothesis_res_list)
write_csv(hmm_hypothesis_res_all, path = here::here("manuscript", "hmm_res.csv"))
```

>>>>

do_pp_checks <- FALSE
hhmm_hypothesis_res_list <- list()

@@

serie_data <- wide %>%
  inner_join(data$patient_data, by = c("hospital_id", "patient_id"), suffix = c("_admission", "")) %>% 
  mutate(.time = day, .serie = patient_id, .observed = breathing_s) %>%
  transform_serie_data()

serie_data_28 <- wide %>%
  right_join(crossing(data$patient_data, day = 0:28), by = c("hospital_id", "patient_id", "day"),  suffix = c("_admission", "")) %>%
  # group_by(patient_id) %>%
  # mutate(
  #   across(
  #     c(starts_with("took_"), all_of("best_supportive_care")),
  #     ~ if_else(is.na(.x) & day > max(day[!is.na(.x)]), .x[which.max(day[!is.na(.x)])], .x)
  #     ),
  # ) %>%
  # ungroup() %>%
  mutate(.time = day, .serie = patient_id, .observed = breathing_s) %>%
  transform_serie_data()

# Extend terminal states and took_XX data for serie_data_28
serie_data_28 <- serie_data_28 %>%
  group_by(.serie) %>%
  mutate(last_day = max(day[!is.na(.observed)]),
         last_state = .observed[day == last_day],
         .observed = if_else(last_state %in% c("Discharged", "Death") & is.na(.observed) & day > last_day,
                             last_state,
                             .observed)) %>%
  mutate(across(c(tidyselect::starts_with("took_"), all_of("best_supportive_care")), ~ if_else(is.na(.x), as.integer(any(.x != 0, na.rm = TRUE)), .x))) %>%
  ungroup()

# Check succesful extension
mismatches <- serie_data_28 %>% inner_join(serie_data, by = c("patient_id", "day")) %>%
  select(-starts_with("."), -last_day,  -last_state) %>%
  mutate(across(everything(), as.character)) %>%
  pivot_longer(!any_of(c("patient_id", "day")), names_to = c("col", "source"), names_sep = "\\.") %>%
  group_by(patient_id, day, col) %>%
  summarise(n_vals = length(unique(value)), .groups = "drop") %>%
  filter(n_vals > 1)

if(nrow(mismatches) > 0) {
  print(mismatches)
  stop("Mismatches in serie_data_28")
}

@@

```{r}
compute_counterfactual_mortality_log_OR <- function(hmm_fit, numerator_serie_data, denominator_serie_data, day = 28) {
  if(nrow(numerator_serie_data) != nrow(denominator_serie_data)) {
    stop("Groups to compare must have equal number of rows")
  }

  numerator_data <- hmm_fit$data
  numerator_data$serie_data = numerator_serie_data

  numerator_state_probs <- posterior_epred_rect(hmm_fit, newdata = numerator_data, method = posterior_epred_state_prob)

  denominator_data <- hmm_fit$data
  denominator_data$serie_data = denominator_serie_data

  denominator_state_probs <- posterior_epred_rect(hmm_fit, newdata = denominator_data, method = posterior_epred_state_prob)

  mortality_state <- which(hmm_fit$data$hidden_state_data$id == "Death")
  samples_probs_numerator <- 
    rowMeans(numerator_state_probs[mortality_state, day + 1, ,])
  samples_probs_denominator <- 
    rowMeans(denominator_state_probs[mortality_state, day + 1, ,])

  log_odds_numerator <- log(samples_probs_numerator) - log1p(-samples_probs_numerator)
  log_odds_denominator <- log(samples_probs_denominator) - log1p(-samples_probs_denominator)

  log_odds_numerator - log_odds_denominator
}

evaluate_treatment_hypothesis <- function(fit, hypothesis, treatment_column, model_subgroup, adjusted, model_check) {
  counterfactual_took <- serie_data_28 %>% mutate({{ treatment_column }} := 1)
  counterfactual_no <- serie_data_28 %>% mutate({{treatment_column }} := 0)

  log_OR_samples <- compute_counterfactual_mortality_log_OR(fit, counterfactual_took, counterfactual_no, day = 28)

  bayesian_hypothesis_res_from_draws(
    draws = log_OR_samples,
    model = paste0("HMM", model_subgroup),
    estimand = "log(OR)",
    hypothesis = hypothesis,
    adjusted = adjusted,
    model_check = model_check
  )
}

evaluate_all_treatment_hypotheses <- function(fit, model_subgroup, adjusted, model_check) {
  rbind(
    evaluate_treatment_hypothesis(fit, hypotheses$hcq_reduces_death, took_hcq, model_subgroup = model_subgroup, adjusted = adjusted, model_check = model_check),
    evaluate_treatment_hypothesis(fit, hypotheses$az_reduces_death, took_az, model_subgroup = model_subgroup, adjusted = adjusted, model_check = model_check),
    evaluate_treatment_hypothesis(fit, hypotheses$favipiravir_reduces_death, took_favipiravir, model_subgroup = model_subgroup, adjusted = adjusted, model_check = model_check)
  )
}

@@

```{r}
hmm_hypothesis_res_list[["trt_group_m1"]] <- evaluate_all_treatment_hypotheses(trt_group_m1_fit, model_subgroup = "simple_group", adjusted ="all treatments, supportive", model_check = "Suspicious")
```

```{r}
if(do_pp_checks) {
  do_common_pp_checks(trt_group_m1_fit)
}
```

@@

```{r}
hhmm_hypothesis_res_list[["trt_group_m2"]] <- evaluate_all_treatment_hypotheses(trt_group_m2_fit, model_subgroup = "simple_group", adjusted ="all treatments, supportive, age, sex", model_check = "Suspicious")
```

@@

```{r}
hhmm_hypothesis_res_list[["trt_group_m3"]] <-  evaluate_all_treatment_hypotheses(trt_group_m3_fit, model_subgroup = "simple_group", adjusted ="all treatments, supportive, age, sex, hospital", model_check = "OK")
```

@@

```{r}
hhmm_hypothesis_res_list[["trt_group_m4"]] <-  evaluate_all_treatment_hypotheses(trt_group_m4_fit, model_subgroup = "simple_group", adjusted ="all treatments, supportive, age, sex, hospital, comorbidities", model_check = "OK")
```

@@

```{r}
hhmm_hypothesis_res_list[["trt_rate_m1"]] <-  evaluate_all_treatment_hypotheses(trt_rate_m1_fit, model_subgroup = "simple_rate", adjusted ="all treatments, supportive", model_check = "Suspicious")
```

@@

```{r}
hhmm_hypothesis_res_list[["trt_rate_m2"]] <-  evaluate_all_treatment_hypotheses(trt_rate_m2_fit, model_subgroup = "simple_rate", adjusted ="all treatments, supportive, age, sex", model_check = "Suspicious")
```

@@

```{r}
hhmm_hypothesis_res_list[["trt_rate_m3"]] <-  evaluate_all_treatment_hypotheses(trt_rate_m3_fit, model_subgroup = "simple_rate", adjusted ="all treatments, supportive, age, sex, hospital", model_check = "OK")
```

@@

```{r}
hhmm_hypothesis_res_list[["trt_rate_m4"]] <- evaluate_all_treatment_hypotheses(trt_rate_m4_fit, model_subgroup = "simple_rate", adjusted ="all treatments, supportive, age, sex, hospital, comorbidities", model_check = "OK")
```

@@

```{r}
draws <- tidybayes::tidy_draws(d_dimer_m1_fit$brmsfit)
hhmm_hypothesis_res_list[["d_dimer_m1_death"]] <-
  bayesian_hypothesis_res_from_draws(
    draws = draws$b_log_peak_d_dimer,
    model = "HMM",
    estimand = "log(HR)",
    hypothesis = hypotheses$d_dimer_increases_death,
    adjusted = "supportive, age, sex, hospital"
  )

@@

```{r}
stop("Calculate hypo")
```

@@

```{r}
stop("Calculate hypo")

draws <- tidybayes::tidy_draws(trt_m7_fit$brmsfit)
hhmm_hypothesis_res_list[["trt_m7_IL_6_death"]] <-
  bayesian_hypothesis_res_from_draws(
    draws = draws$b_log_peak_IL_6,
    model = "HMM",
    estimand = "log(HR)",
    hypothesis = hypotheses$IL_6_increases_death,
    adjusted = "age, sex, hospital"
  )

@@

```{r}
stop("Calculate hypo")

@@

```{r}
hhmm_hypothesis_res_list[["c_trt_rate_m1"]] <- evaluate_all_treatment_hypotheses(c_trt_m1_fit, model_subgroup = "complex_group", adjusted ="all treatments, supportive", model_check = "Suspicious")
```

```{r}
if(do_pp_checks) {
  do_common_pp_checks(c_trt_m1_fit)
}
```

@@

```{r}
hhmm_hypothesis_res_list[["c_trt_rate_m2"]] <- evaluate_all_treatment_hypotheses(c_trt_m2_fit, model_subgroup = "complex_group", adjusted ="all treatments, supportive, age, sex", model_check = "Suspicious")

@@

```{r}
hhmm_hypothesis_res_list[["c_trt_rate_m3"]] <- evaluate_all_treatment_hypotheses(c_trt_m3_fit, model_subgroup = "complex_group", adjusted ="all treatments, supportive, age, sex, hospital", model_check = "Suspicious")

@@

```{r}
hhmm_hypothesis_res_list[["c_trt_rate_m4"]] <- evaluate_all_treatment_hypotheses(c_trt_m4_fit, model_subgroup = "complex_group", adjusted ="all treatments, supportive, age, sex, hospital, comorbidities", model_check = "Suspicious")

@@

```{r}
hmm_hypothesis_res_all <- do.call(rbind, hhmm_hypothesis_res_list)
write_csv(hmm_hypothesis_res_all, path = here::here("manuscript", "hmm_res.csv"))
```
